Building socket-aware BPF programs

### 2018-10-27 paper/paper.tex
paper: Improve general structure and wording

As with many other structure types, Linux sockets are reference-counted to
track usage and handle memory release. In general, to ensure the safety of
socket for the duration of access in BPF programs, when sockets are referenced
from BPF programs the reference count for the socket needs to be incremented.
Two options were considered for how to guarantee this: Implicit reference
tracking and explicit reference tracking.

Implicit reference tracking calls for the BPF infrastructure to handle
reference tracking, and to hide this detail from the BPF API. Whenever a socket
lookup function is called, a reference is taken and the socket is added to a
reference list. At the end of the BPF program execution, the core code could
walk this reference list and release each reference. However, collecting
references and releasing them at the end of the BPF program invocation has the
unfortunate overhead that even the execution of BPF programs that do not make
use of the new socket lookup helpers would need to pay the cost of checking the
list of references to free. When dealing with sufficiently demanding use cases
such as those which use XDP, even the few instructions required to implement
such a check may have a measurable impact on performance.

Explicit reference tracking calls for reference acquisition and release
semantics to be built into the API. If references must be taken, then requiring
BPF program writers to explicitly handle these references ensures that the
program writer must understand the potential impact this may have on the
operation of the program (including atomic operation cost), and it also ensures
that the cost is only borne by programs that make use of this feature.

\todo{Lead into the implementation} This safety is provided by tracking pointer
values through each conditional path in the program, and rejecting the load of
programs that fail to "balance" resource acquisition (lookup) with release.

Calls to helper functions which acquire references to resources are annotated
in the verifier to associate the acquisition and release functions. When the
instruction that acquires a resource is processed, the verifier state is
updated to include a resource identifier for the resulting value, and the
register's value is associated with this identifier. When processing a
corresponding release helper call, if a parameter to the function contains a
resource identifier, then the corresponding identifier is removed from the
verifier state. If any paths in the program reach the final (\verb+BPF_EXIT+)
instruction and the verifier state still contains resource identifiers, then
the program is considered unsafe as it has leaked the resource.

@@

The following considerations were made for implementing the lookup helpers:

@@

The consideration of each of these items reflects the definition of the lookup
helpers (Listing~\ref{lst:helper-api}):

@@

@@

The implementation intruduced in the 4.20 release cycle includes support for
looking up TCP and UDP sockets for IPv4 and IPv6 traffic, using the \verb+tc+
and \verb+sk_skb+ hook points. A proposed patchset extends this to allow the
same helpers to be used from the \verb+xdp+ hook points.

>>>>

As with many other structure types, Linux sockets are reference-counted to
track usage and handle memory release. In general to ensure the safety of
socket for the duration of access in BPF programs, this reference count needs
to be incremented when sockets are referenced from BPF programs. Two options
were considered for how to guarantee this: Implicit reference tracking and
explicit reference tracking.

\subsubsection{Implicit reference tracking.}

This calls for the BPF infrastructure to handle reference tracking, and to hide
this detail from the BPF API. Whenever a socket lookup function is called, a
reference is taken and the socket is added to a reference list. At the end of
the BPF program execution, the core code could walk this reference list and
release each reference. However, collecting references and releasing them at
the end of the BPF program invocation has the unfortunate overhead that even
the execution of BPF programs that do not make use of the new socket lookup
helpers would need to pay the cost of checking the list of references to free.
When dealing with sufficiently demanding use cases such as those which use XDP,
even the few instructions required to implement such a check may have a
measurable impact on performance.

\subsubsection{Explicit reference tracking.}

This calls for reference acquisition and release semantics to be built into the
API. If references must be taken, then requiring BPF program writers to
explicitly handle these references ensures that the program writer must
understand the potential impact this may have on the operation of the program
(including atomic operation cost), and it also ensures that the cost is only
borne by programs that make use of this feature.

\subsubsection{Implementation.}

Based upon the tradeoffs described above, explicit reference tracking was
chosen. From an implementation perspective, this requires tracking pointer
values through each conditional path in the program, and rejecting the load of
programs that fail to "balance" resource acquisition (lookup) with release.
This implementation works as follows.

Calls to helper functions which acquire references to resources are annotated
in the verifier to associate the acquisition and release functions. When the
instruction that acquires a resource is processed, a resource identifier is
allocated for this resource. This identifier is kept in a list in the verifier
state, and it is also associated with the register which receives the resulting
resource pointer. When processing a corresponding release helper call, if a
parameter to the function contains a register that is associated with a
resource identifier, then the identifier is removed from the verifier state. If
any paths in the program reach the final (\verb+BPF_EXIT+) instruction and the
verifier state contains any resource identifier association, then the program
is considered unsafe as it has leaked the resource.

Some instruction calls are restricted while holding a resource reference to
avoid leaking references, for example the \verb+bpf_tail_call+ helper function.

@@

The following considerations were made for implementing the lookup helpers. The
consideration of each of these items reflects the definition of the lookup
helpers in Listing~\ref{lst:helper-api}.

@@

@@

Multiple optimizations were proposed during development of this feature to
reduce the runtime overhead of using the socket lookup helpers. Two such
optimizations are described below: Skipping reference counting when unnecessary
to save on atomic instructions, and allowing the use of direct packet pointers
for the tuple.

@@

The implementation intruduced in the 4.20 release cycle includes support for
looking up TCP and UDP sockets for IPv4 and IPv6 traffic, using the
\verb+SCHED_CLS+, \verb+SCHED_ACT+ and \verb+SK_SKB+ hook points. A proposed
patchset extends this to allow the same helpers to be used from the \verb+XDP+
hook points~\cite{sk-lookup-xdp}.
